---
title: "inference.Rmd"
author: "Damian"
date: "May 30, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Vitamin C and Tooth Growth in Guinea Pigs*
===============================================================================================
load the data
```{r}
library(ggplot2)
library(datasets)
data("ToothGrowth")
```
## Exploration
I crossed the dose and supplement factors, and plotted tooth length for each level
```{r}
ToothGrowth$DS <- interaction(ToothGrowth$supp, ToothGrowth$dose)
ggplot(data = ToothGrowth, aes(y=len, x=sort(DS), fill = supp)) + geom_boxplot() + scale_x_discrete(labels=c("0.5","1.0","2.0","0.5","1.0","2.0")) + labs(x="Dose", title="Effect of Dose and Supplement on Length")
```
## Analysis

I wrote a function that would do a two-sample t-test on pairs of tooth length vectors  
the null hypothesis for these tests will be that the difference of the means is 0 
```{r}
do_t_tests <- function(a_list,alt_hyp){
        # performs two-sample t-tests on all list items
        for(group in seq_along(1:(length(a_list) - 1))){
                for(other_group in (group + 1):length(a_list)){
                    t_test <- t.test(a_list[[other_group]],a_list[[group]],alternative=alt_hyp)    
                    t_tests[[paste((names(a_list))[other_group]," vs. ",names(a_list)[group])]] <<- t_test
        }
        }
}
```

I made a list of 6 tooth-length vector, each for one dose-supplement factor level
```{r}
supp_dose_combos <- list()
for(supp_dose_pair in levels(ToothGrowth$DS)){
        supp_dose_combos[[supp_dose_pair]] <- ToothGrowth[(ToothGrowth$DS == supp_dose_pair),]$len
}
```

I split the 6 vectors into 2 groups   
* one for guinea pigs that got OJ  
* one for gp's that got VC
```{r}
OJ_groups <- supp_dose_combos[grepl(pattern="OJ",x=names(supp_dose_combos))]
VC_groups <- supp_dose_combos[grepl(pattern="VC",x=names(supp_dose_combos))]

supp_group <- list(OJ_groups=OJ_groups, VC_groups=VC_groups)
```
I also split the 6 vectors into 3 groups   
one for each dose level: 0.5, 1, and 2
```{r}
half_groups <- supp_dose_combos[grepl(pattern="0.5",x=names(supp_dose_combos))]
one_groups <- supp_dose_combos[grepl(pattern="1",x=names(supp_dose_combos))]
two_groups <- supp_dose_combos[grepl(pattern="2",x=names(supp_dose_combos))]

dose_group <- list(half_groups=half_groups,one_groups=one_groups,two_groups=two_groups)
```
I performed two-sample t-tests on all samples that had the same level for exactly one factor 
```{r}
t_tests <- list()

for(each_group in supp_group){
    do_t_tests(each_group,alt_hyp = "greater")
}
for(each_group in dose_group){
    do_t_tests(each_group,alt_hyp = "two.sided")
}
```
I displayed preliminary results  
if 0 did not fall in the 95% confidence interval...    
then the groups were said to be significantly different
```{r}
for(ind in 1:length(t_tests)) {ifelse(sum(0 > t_tests[[ind]]$conf.int)==0 | sum(0 > t_tests[[ind]]$conf.int)==2 , print(paste(names(t_tests)[[ind]],"are significantly different")) ,print(paste(names(t_tests)[[ind]],"are not significantly different")))}
```
I applied the Benjamini-Hochberg correction to fix the FDR at 5%
```{r}
p_scores <- numeric()
for(t_tesst in t_tests){p_scores <<- c(p_scores,t_tesst$p.value)}
adjusted_p_scores <- p.adjust(p_scores,method="BH")
for(ind in 1:length(t_tests)) {ifelse(adjusted_p_scores[ind] < .05 , print(paste(names(t_tests)[[ind]],"are significantly different")) ,print(paste(names(t_tests)[[ind]],"are not significantly different")))}
```

I plotted the p-scores of two sample t-tests on samples with fixed dose
```{r}
the_names <- names(t_tests)

Supp_P <- adjusted_p_scores[7:9]
Supp_name <- the_names[7:9]


sup_df <- data.frame(Supp_P,Supp_name)


par(mfrow=c(1,1))
with(sup_df,plot(Supp_P~Supp_name,main="probability of no significant difference between supplement effectiveness at different doses"))
```

    
## Conclusion
* it appears that larger vitamin C doses significantly increase tooth growth in guinea pigs
* OJ is a more effective supplement at low doses 
* at high doses, neither supplement is better
Assumptions
* all environmental factors were kept constant across groups
* all guinea pigs were from the same population
* the GP's in the study are representative of the GP population 
