Steps Assignment
========================================================

## Part 0 : load some packages and get the raw data into a dataframe

```{r}
library(ggplot2)
library(plyr)
library(dplyr)

download.file(url="https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
              destfile = "raw_steps_data.zip")

unzip("raw_steps_data.zip",exdir = "steps_assignment")

raw_steps_data <- read.csv("steps_assignment/activity.csv")
```

## Part 1 : Steps by Day

### group by date and sum the steps
```{r}
steps_per_day <- group_by(raw_steps_data,date) %>% summarise(Steps = sum(steps,na.rm=TRUE)) 
```

### plot the histogram
```{r}
qplot(x=steps_per_day$Steps,binwidth=2000,ylab="frequency", xlab = "steps per day", main="absolute frequency of steps per day, bin size = 2000")
```

### get the mean and median for steps-per-day
```{r}
SpD_median <- median(steps_per_day$Steps,na.rm = TRUE)
SpD_mean <- mean(steps_per_day$Steps,na.rm=TRUE)
```
The mean is `r SpD_mean`.
The median is `r SpD_median`.

## Part 2 : Steps by Time

### group by time intervals and calculate the mean number of steps
```{r}
interval_steps <- group_by(raw_steps_data,interval) %>% 
    summarize(mean_steps = mean(steps,na.rm =TRUE))
```

### plot a timeseries of average step-count
```{r}
qplot(data = interval_steps, x= interval, y = mean_steps, geom="line",main="step-count throughout the day")
```

### find most active interval
```{r}
most_active_steps <- max(interval_steps$mean_steps)
most_active_interval <-            interval_steps[[match(most_active_steps,interval_steps$mean_steps),1]]
```
The most active 5 minute was at `r most_active_interval`

## Part 3 : Impute missing data and remake steps per day histogram

## impute missing data by replacing each NA with the average of its interval
```{r}
imputed_steps_data <- raw_steps_data
for(entry in 1:nrow(imputed_steps_data)){
    
    if(is.na(imputed_steps_data[["steps"]][entry])){
            the_interval <- imputed_steps_data[["interval"]][entry]
            sub_value <- interval_steps[interval_steps$interval==the_interval,][["mean_steps"]]
            imputed_steps_data[["steps"]][entry] <- sub_value            
            
    }
}
```

## recalculate total steps for each day
```{r}
new_steps_per_day <- group_by(imputed_steps_data,date) %>% summarize(daily_steps = sum(steps))
```

## replot histogram
```{r}
qplot(x=new_steps_per_day$daily_steps,binwidth=2000,ylab="frequency", xlab = "steps per day",main="absolute frequency of steps per day, bin size = 2000")
```

## recalculate mean and median
```{r}
new_median <- median(new_steps_per_day$daily_steps)
new_mean <- mean(new_steps_per_day$daily_steps)
```
The mean of imputed data is `r new_mean`, and the median is `r new_median`

### Part 4 : plot time-series of the average weekday and weekend

## convert the date variable into a factor with two levels: weekday and weekend
```{r}
imputed_steps_data <- mutate(imputed_steps_data, DOW = weekdays(as.Date(date)))
imputed_steps_data <- mutate(imputed_steps_data, week_end_day = ifelse(DOW=="Saturday" | DOW == "Sunday", "Weekend","Weekday"))
imputed_steps_data <- mutate(imputed_steps_data, week_end_day = factor(week_end_day))
```

## group by weekend vs. weekday plus time interval, get mean steps for each group
```{r}
group_day_int <- group_by(imputed_steps_data,week_end_day,interval) %>%
                                        summarize(the_mean=mean(steps))
```

## plot time-series of steps for weekends and weekdays
```{r}
qplot(x=interval,y=the_mean,facets = .~week_end_day,geom="line",data=group_day_int)
```
