---
title: "Regression_Cars"
author: "Damian"
date: "June 11, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(datasets)
data(mtcars)
```

### does transmission type have an effect on mpg?  
We will answer this question!

### exploratory analysis
```{r}
ggplot(data = mtcars, aes(x = factor(am), y = mpg, color = am)) + geom_point()
```

Seems like cars with manual transmission have higher mpg

```{r}
auto_mean <- mean(mtcars$mpg[mtcars$am == 0])
man_mean <- mean(mtcars$mpg[mtcars$am == 1])
```

Mean mpg by transmission type
Automatic : $`r auto_mean`$
Manual : $`r man_mean`$


### Linear Models
I fit three models
* One predicting mpg only using transmission type
* One predicting mpg only using weight
* One predicting mpg using weight and transmission

# transmission type alone
```{r}
model1 <- lm(data = mtcars, formula = mpg~am)
summary(model1)
```
# transmission and weight
```{r}
model2 <- lm(data = mtcars, formula = mpg ~ am + wt)
summary(model2)
```

# weight alone
```{r}
model3 <- lm(data = mtcars, formula = mpg ~ wt)
summary(model3)
```

### What does it mean?
```{r}
p_am <- summary(model1)$coefficients[2,4]
p_am_wt <- summary(model2)$coefficients[2,4]
```
Transmission type is a statistically significant predictor when used alone.
The p-value : $`r p_am`$ can be interpreted as saying that there's a 2.8% chance of 
observing a correlation of this magnitude if there was no relationship between mpg
and transmission type.

However, when weight is included in the model, transmission type is far from predictive, with p-value equal to $`r p_am_wt`$.


### residual plot
```{r}
residual_df <- data.frame(row.names=names(model3$residuals), resid = model3$residuals)
residual_df$AM <- mtcars$am
ggplot(data = residual_df, aes(x=resid, fill = as.factor(AM))) + geom_histogram(binwidth=1.5)
```

### leverage
```{r}
plot(hat(model.matrix(model2)))
```

