---
title: "Storm_HW"
author: "Damian"
date: "May 25, 2016"
output: html_document
---

**Most badass kinds of storms**
============================================================================

## Synopsis
* The independent variable was storm type, which was coded by EVTYPE
* The dependent variables were negative effects, coded by FATALITIES, INJURIES, PROPDMG, CROPDMG
* The data were grouped by storm type, and the mean of each negative effect was taken.
* The metrics were chosen: one for economic effects, the other for human effects.
* The data were arranged into two lists by the different metrics. 
* The three most destructive storms by each metric were prioritized for research into preventetive measures


### download file and packages and load into data frame
```{r}
cat("in chunk 1")
library(plyr)
library(dplyr)
library(ggplot2)
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2",destfile = "stormz.bz2")
stormz <- read.csv("stormz.bz2")
cat("leaving chunk 1")
```

## Data Processing

### group by storm type and sum over each kind of negative effects
```{r}
cat("in chunk 2")
type_storms <- group_by(stormz, EVTYPE)
type_storms_sum <- summarize(type_storms, 
                             fatal_avg = mean(FATALITIES, na.rm = TRUE),
                             injur_avg = mean(INJURIES, na.rm = TRUE),
                             propdmg_avg = mean(PROPDMG, na.rm = TRUE),
                             cropdmg_avg = mean(CROPDMG, na.rm = TRUE))
```

### choose arbitrary health metric: 1 death = 5 injuries
```{r}
cat("in chunk 3")
fat_inj <- mutate(type_storms_sum, health_damage = fatal_avg * 5 + injur_avg)
fat_inj <- arrange(fat_inj, desc(health_damage))
```

### choose arbitrary econ metric : 1 property = 1 crop
```{r}
cat("in chunk 4")
crop_prop <- mutate(type_storms_sum, crop_prop_damage = propdmg_avg + cropdmg_avg)
crop_prop <- arrange(crop_prop, desc(crop_prop_damage))
```
### note that the health action is all in the fourth quantile
```{r}
cat("in chunk 5")
health_quant <- cut(fat_inj$health_damage,quantile(fat_inj$health_damage,probs=c(.75,1.0)))
fat_inj <- mutate(fat_inj,split_quantiles = health_quant)
health_fourthQ <- group_by(fat_inj,split_quantiles) %>% summarize(mean(health_damage))
mean_health_dmg <- health_fourthQ[[1,2]]
```
### likewise for econ
```{r}
cat("in chunk 6")
econ_quant <- cut(crop_prop$crop_prop_damage,quantile(crop_prop$crop_prop_damage,probs=c(.75,1.0)));
crop_prop <- mutate(crop_prop,split_quantiles = econ_quant);
crop_prop_fourthQ <- group_by(crop_prop,split_quantiles) %>% summarize(mean(crop_prop_damage));
mean_crop_prop_dmg <- crop_prop_fourthQ[[1,2]]
```

### construct mini-dataframes for health and econ's most destructive storms
```{r}
cat("in chunk 7")
mini_crop_prop <- select(crop_prop, EVTYPE, crop_prop_damage)[2:4,]
mean_crop_prop <- data.frame(EVTYPE=factor("mean"), crop_prop_damage = mean_crop_prop_dmg)
mini_crop_prop <- rbind(mini_crop_prop,mean_crop_prop)
mini_crop_prop <- mutate(mini_crop_prop, EVTYPE = as.character(EVTYPE))

mini_fat_inj <- select(fat_inj, EVTYPE, health_damage)[c(1,3,4),]
mean_fat_inj <- data.frame(EVTYPE = factor("mean"), health_damage = mean_health_dmg)
mini_fat_inj <- rbind(mini_fat_inj, mean_fat_inj)
mini_fat_inj <- mutate(mini_fat_inj, EVTYPE = as.character(EVTYPE))
```
## Results

### plot the top 3 sources of damage against the median for the fourth quartile
```{r}
cat("in chunk 8")
par(mfcol=c(1,2), mar = c(4,4,3,1))
with(mini_crop_prop, barplot(height = crop_prop_damage,names.arg = EVTYPE, xlab= "storm type", ylab="economic damage",main="mean economic damage caused by most destructive storms"))
with(mini_fat_inj, barplot(height = health_damage,names.arg = EVTYPE, xlab= "storm type", ylab="health damage",main="mean health damage caused by most destructive storms"))
```
*here we show the damage caused by the three most destructive storm types and the mean damage caused by fourth quantile storm types.*  

*the left bar-plot investigates economic damage, while the right bar-plot investigates human damage*

### Conclusion
**Wind, extreme temperatures, and non-liquid precipitation are the most dangerous weather events**
**Water related events are most destructive**


